%I want to express my great appreciation to several people.

I am honored to have the privilege of continuing my education and being part of the science community. Pursuing this journey has been a goal of mine for many years, and it would not have been possible without the support of many people along the way.

First, I would like to thank the graduate committee at Baylor University for giving me the opportunity to be part of their program. This experience had a significant impact on my personal and professional life. I would like to especially thank Dr. Mathews, Dr. Hilton, and my advisor Dr. Hatakeyama for their continued support and inspiration.

Second, I would like to thank my family who always supported me in following my dreams, and my friends who made living abroad tolerable and full of fun memories. I would like to especially thank the Manzano family for taking care of me, and my best friends Shabnam and Ahlam for always being by my side.

Finally, I am grateful for the financial support I received from the Saudi Arabian Cultural Mission (SACM) which made my education possible.   

%% outline %%
%I am honored to having the privilege of doing science/ research for living. to contrinute to the process.
%reaching this point in my career would not have been possible with out the help and support of many people.
%I would like to thing my family. 
%I would like to thank baylor univeristy.
%I would like to thank HEP gorup.  
%I would like to thank my friends. 
%Finally, I am grateful for (SACM) for the financial support during the masters program. 
