% PF ECAL Cluster Calibration using Boosted Decision Tree %

% Notes: source PF paper
%.......................

% 3.5 Calorimeter cluster calibration:
%.....................................
%% mentioned in ch2. In the reconstruction algorithm (PF), photons & neutral hadrons are reconstructed from the calorimeter clusters.
%% *reminder: in we identify PF candidates in order: Mu => electrons, isolated (photons) => nonisolated photons (from pi), hadrons (charged, neutral)
%% *isolated meaning (not in part of jet)  
%% this identified photons & hadrons are calibrated.
%% the estimated true energy of each particle is taken to be the raw calorimeteric energy. E(Ecal) photons.  



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%  
\section{Boosted Decision Tree}
%write a simple introduction to explain how BDT works and how it helps in PF cluster regression to  correct the PF cluster energies. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ECAL Cluster Calibration Using BDT}
% notes slide 24 source: https://www.desy.de/~garutti/LECTURES/ParticleDetectorSS12/L10_Calorimetry.pdf %
% To make a statement about the energy of a particle we need to calcluate two things %
% 1) Detector response: relationship between measured signal and deposited energy
% 2) Detector resolution: precision with which the unknown energy can be measured
%% slide 25 Response and linearity %%
% “response = average signal per unit of deposited energy”
% A linear calorimeter has a constant response, Electromagnetic calorimeters are linear %
% Hadronic calorimeters are not linear % 

PF Cluster regression is used to correct PF cluster energies lost due to tracker material, gaps, dead channels etc. The ML method implemented is a Boosted Decision Tree (BDT) based semi-parametric regression. To estimate the correction to the PF ECAL Cluster, we consider the case where a one photon has deposited all its energy in one PF cluster in the ECAL. This will allow us to calculate the correction factor = E “gen photon” / E “raw pf cluster”. 

\subsection{Strategy}
%\subsection{Plan for Run3 (2024) Calibration}

We are adopting the same variables used previously in Run 2 & Run 3 (2023). The PF cluster regression is done intro two steps:  

1) Training (from where we will get trained regression model). 

2) The validation of the training is done on NOPU and PU Zero material double photon samples.

\subsection{Samples used and Training Details}

Training is done on no PU Zero material double photon particle gun. During training we are going to adopt the same variables used in Run 2 calibration for Run 3 (2024). 

The training input variables include independent variables or features used in the training:  
	clusrawE , ieta, iphi (EB) , ix, iy (EE), ietamod20, iphimod20 (EB only), (clusPS1+clusPS2)/clusrawE (EE only), number of hits in the cluster (which takes values 1, 2 and 3 - it takes 3 if nhits >= 3). The used  

Target: log(Egen/Eraw). 

 
\subsection{Validation}

The validation is done through few steps. First from the training model we get the value of the correction factor, then calculated:  the corrected cluster energy = E “raw pf cluster” * correction factor Plot: the response: E “corrected PF cluster” / E “gen photon” Fit the plot with: double CB function. From the fitted curve we find: mean, effective sigma. We can also get the mean, effective sigma for raw PF cluster energy “with no correction” to compare. 

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Discussion}

we can validate the training by comparing the results of the new corrected ECAL cluster in 133X (blue line), to the current  correction in 126X (green) and  raw PF ECAL cluster (red).
generally we see that the new  correction very close to the current used calibration

Usually, the since the PF clustering is performed separately in each region of the ECAL: ECAL Barrel (EB), ECAL Endcaps (EE).
Presenting the results in a similar way.

Overview of the results: we checked the 2024 double photon calibration samples for PF ECAL clusters. In general, the existing calibration derived from 2022 samples seems to continue to be working well. 

 
\subsection{ECAL Barrel}
first starting with the NOPU sample then the PU case.
plots show response (resolution) vs Pt gen in GeV and in their corresponding eta range.
\include{PU_EB_FULL_plots}
\include{PU_EB_ZS_plots}
\include{NOPU_EB_FULL_plots}
\include{NOPU_EB_ZS_plots}

\subsection{ECAL Endcap}
in EE region:
\include{NOPU_EE_FULL_plots}
\include{NOPU_EE_ZS_plots}
\include{PU_EE_FULL_plots}
\include{PU_EE_ZS_plots}



HLT vs offline PF ECAL cluster
first for NoPU samples
\include{NoPU_HLT_offline_plots}
Second for PUU samples
\include{PU_HLT_offline_plots}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\section{Citations}
%Here is a citation \cite{fake1}, and here is another \cite{fake2}. Citations are
%nice. Depending on your choice of bibliography, there may be different formats
%you can use. For example, Chicago provides a family of short citation commands.BB
