% intro CMS %

CMS detector is one of the main detectors at the LHC. It is a general-purpose detector and its physics program rang from studying the Standard Model
(including the Higgs boson) to searching for extra dimensions and particles that could make up dark matter.
CMS stands for Compact Muon Solenoid (CMS). (CMS is designed to detect muons very accurately, and it has the most powerful solenoid magnet ever made.
This detector weights 14,000-tonne but it is quite compact for all the detector material it contains. It is at 15 m high and 21 m long, fix me).
The design of CMS and its subdetectors are based on the geometry of its solenoid magnet. Inside the CMS starting from collision point we have: Silicon tracker designed to minimally interact with particles, electromagnetic and hadronic calorimeters will stop and absorb the particles completely.
Outside the magnet there are some of HCAL subdetectors and muon systems.

\subsection{Superconducting Magnet} 
It is superconducting magnet that produces 3.8 T magnetic field inside a solenoid and quickly drops off outside it. This magnetic field is used to determine the momenta of charged particles since the trajectories of the particles bend in the field.

\subsection{Inner Tracker} 
The first subdetector particles created from the collision will interact with after leaving the beam pipe is the tracker. The most inner parts of the CMS detector are: the pixels, silicon microstrip detectors that surrounds the pixels. The main function of the tracker is to measure the curvature of charged particles with pt > 1 GeV. From which we can use to determine the momentum. When a charged particle goes through the tracker it will leave a hit (tiny signal that will later be amplified and detected) in the tracker layers. These hits could be aligned to get the trajectory of the particle. The signals will be stored for microseconds and then processed before being sent to a laser to be converted to an infrared pulse Which will be transmitted for analysis.

The pixel detector: It is also the closest detector to the beam pipe. It has 4 cylindrical layers at (3,7,11,16 cm) . Each of the four layers is composed of individual silicon module (? How many), divided into little silicon sensors (pixels).  Each of these silicon pixels is 100µm by 150µm.

The silicon strips: after the pixels the particles go through the silicon strip detector, which has10 layers. (130 cm~ 1.3m). 4 inner barrel layers (TIB) with two inner endcaps (TID) each has 3 small disks. Then we have 6 outer barrel layer (TOB) surrounds both TIB, TID with two endcaps (TED) closing off the tracker. this part contains 15,200 highly sensitive modules with a total of about 10 milli on detector strips. (insert real/sketch updated picture of each subdetector) 

\subsection{Electromagnetic Calorimeter}
It is a homogeneous calorimeter made of lead tungstate crystals. Measures the energy of electrons and photons. It is made up of a barrel section and two endcaps, that would form a layer between the tracker and the HCAL.

Tungstate crystals are heavier than stainless steel but it is transparent. It scintillates when electrons and photons pass through it. Which will produce light that is that is proportion to the particle energy. Glued onto the back of each of the crystals a Photodetectors, which have been especially designed to work within the high magnetic field, will detect the scintillation light and convert it to an electrical signal that is amplified and sent for analysis.

The barrel region: consists of 61,200 crystals formed into 36 “supermodules”, each weighing around three tones and containing 1700 crystals. Here each crystal (2.2 x 2.2 x 23 cm). it can contain more than 98 of the electrons and photons energy. The endcap regions close off the barrel at each end and are made up of almost 15,000 more crystals. Each crystal (3 x 3 x 22 cm) which corresponding to 24.7 radiation lengths. For more precision, the ECAL also contains a pre shower detector that is Finer grained detector and located before either endcap disk has two layers, each layer is made of two orthogonal layers of silicon sensors, interspersed with lead layers that serve to generate electromagnetic showers. This detector serves two goals Find the photons decaying form a neutral pion to discriminate them from prompt photons.  Indicate the presence of an electron or a photon in the ECAL be requiring a related signal in the pre shower. (what pics to include?) 

\subsection{Hadronic Calorimeter}
HCAL measures directly the energy of hadrons (such as neutrons, Pions, and kaons) and indirect the presence of non-interacting, uncharged particles such as neutrinos by seeing the imbalance in the momentum and energy (measured in the sideways transverse direction relative to the beam line).

HCAL is hermetic and sampling calorimeter. Hermetic makes sure it captures, to the extent, every particle emerging from the collisions. Sampling calorimeter meaning is made of alternating layers of “absorber” (dense material - brass or steel) and fluorescent “scintillator” (tiles of plastic) that produce a rapid light pulse when the particle passes through.

The HCAL is organized into:  
Barrel region: (HB) these form the last layer of detector inside the magnet coil while a few additional layers the outer barrel (HO), sit outside the coil, ensuring no energy leaks out the back of the HB undetected.  
Endcap regions: measure particle energies as they emerge through the ends of the solenoid magnet. Forward sections positioned at either end of CMS; to pick up the myriad particles coming out of the collisionregion at shallow angles relative to the beam line, these receive the bulk of the particle energy contained in the collision, so they must be very resistant to radiation and use different materials to the other parts of the HCAL.

When a hadronic particle hits a plate of absorber, in this case brass or steel, an interaction can occur producing numerous secondary particles. As these secondary particles flow through successive layers of absorber.
they too can interact and a cascade (shower) of particles results.

As this shower develops, the particles pass through the alternating layers of active scintillation material causing them to emit blue-violet light. Within each tile tiny optical absorb this light. clear optic cables then carry the light a way to readout boxes located at locations within the HCAL volume. signals from successive tiles, one behind the other, are added optically to form towers. These summed optical signals are converted into fast electronic signals by photos sensors. 
(insert sketch that shows the subdetectors of the HCAL)

\subsection{Muon Detector}
Detecting muons is one of CMS’s most important tasks. Unlike most particles muons are not stopped by any of CMS’s calorimeters. Because of that the chambers to detect muons are placed at the very edge of the experiment where they are the only particles likely to register a signal. Muon chambers that include: Drift tubes (DTs) and cathode strip chambers (RPCs) are arranged in concentric cylinders around the beam line in the barrel region. In endcap region, we have CSCs and resistive plate chambers (RPCs) which make up the “endcaps” disks that cover the barrel (add gas electron multiplier chambers (GEMs).

The muon is measured by fitting a curve to hits among the four muon stations (MS) which sit outside the magnet coil and are interleaved with iron return yoke plates. (Insert picture that shows the placement of each type in both regions). 
% figures % 
